Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-06-24 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

From 1989 (2-3 years after disturbance) to 2015 (28-29
years after disturbance), 828388 individual trees and 591 botanical species
pertaining to 223 genus and 64 botanical families were recorded.
For undisturbed plots Richness, Shannon and Simpson diversities remained
stable over the 30 years monitored. In disturbed communities the
taxonomic richness increased after low disturbance intensity, reaching a
maximum gain of 14 botanical genera (plot 3 from treatment 2) while it
followed unimodal trajectories after intense disturbance, decreasing for
ten years before recovering to pre-disturbance values. In all disturbed
plots the taxonomic evenness (Shannon and Simpson diversities)
increased, following unimodal trajectories with a maximum, reached after
around 20 years, positively correlated to the disturbance treatment
(\(\rho_{spearman}^{Shannon}=0.86\), and
\(\rho_{spearman}^{Simpson}=0.89\)). Return towards initial evenness
values was beginning after 30 years except for two T3 plots (plots 8 and
12) which evenness still increased, suggesting similar but delayed
trajectories \ref{fig:IDHplot}.

@@

\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3. Lower panels, Relationship between the initial \%AGB removed and the median values of Simpson \textbf{(c)} and Rao \textbf{(d)} diversities at three times after disturbance. Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.}\label{fig:IDHplot}
\end{figure*}

@@

The trajectories of taxonomic and functional composition were visualized
in a two dimensional ordination space mapping the successive inventories
according to their flora and corresponding traits. Classifications were
performed using either abundance-based Bray-Curtis (Figure
\ref{fig:NMDSplans}) or incidence-based Jaccard dissimilarity, both
giving similar results only analysis using Bray-Curtis dissimilarity are
presented.

While both taxonomic and functional composition remained stable in
undisturbed communities, they followed consistent trajectories over time
after disturbance which revealed significant compositional changes.
According to the mapping of functional traits (see appendix I) these
compositional changes corresponded to shifts towards species with more
acquisitive functional strategies, from communities with high average WD
to high average SLA and chlorophyll content. For disturbed communities
the distance of successive inventories to the 1989 reference inventory
followed unimodal trajectories translating cyclic compositional changes
with a recovery of the initial composition (Figure \ref{fig:NMDSplans}).
The maximum dissimilarity with the initial state was positively
correlated to the disturbance treatment for both taxonomic and
functional composition (\(\rho_{spearman}^{taxonomic}=0.91\) and
\(\rho_{spearman}^{functional}=0.96\) respectively) and the time at
maximum was reached around 26 years after disturbance for taxonomic
composition and 22 years for functional composition.

@@

\caption{Trajectories of the plots in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) regarding the 6 leaf and stem functional traits, the maximum allometric height and seed mass class. Plots trajectories are first represented in the two-dimensional space from the NMDS performed for the 30 years after disturbance based on Bray-Curtis dissimilarity measures between successive inventories (Upper panels \textbf{(a)} and \textbf{(b)}). Then the lower panels (\textbf{(c)} and \textbf{(d)}) represent the euclidean distance to initial condition along the 30 sampled years. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3). The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.}\label{fig:NMDSplans}
\end{figure*}

@@

The changes observed in plots functional composition went hand to hand
with consistent trajectories of the 8 functional traits
visualized with the trajectories of community weighted means (CWM) of
leaves economics (leaves thickness, chlorophyll content, toughness and
specific area), wood economics (wood specific gravity, bark thickness),
and life history traits (seed mass and maximum adult height) (Figure
\ref{fig:CWM}).

Except for leaf chlorophyll content, which continued to increase for
some T3 and T2 plots 30 years after disturbance, all traits and seed
mass proportions followed unimodal trajectories either stabilizing or
returning towards their initial values. Thirty years after disturbance
the weighted means of communities specific maximum height at adult stage
(\emph{Hmax}), leaf toughness (\emph{L\_toughness}) and wood specific
gravity (\emph{WD}) remained significantly lower than their initial
value (Figure \ref{fig:CWM}). The weighted means of bark thickness
(\emph{Bark\_thick}) similarly remained substantially higher than
initially for all disturbed plots while the specfic leaf area
(\emph{SLA}) had almost recovered its initial value. For all traits the
maximum difference to initial state was correlated to the disturbance
intensity (\(\rho_{spearman}^{L_{thickness}}=0.67\),
\(\rho_{spearman}^{L_{chloro}}=0.45\),
\(\rho_{spearman}^{L_{toughness}}=-0.43\),
\(\rho_{spearman}^{SLA}=0.93\), \(\rho_{spearman}^{WD}=-0.78\),
\(\rho_{spearman}^{Bark-thickness}=0.88\),
\(\rho_{spearman}^{Hmax}=-0.48\)).

@@

\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}). Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:CWM}
\end{figure*}

@@

Communities functional redundancy remained stable in control
plots but after disturbance the redundancy trajectories were quite
variable (See appendix I) and apparently independently of the initial
disturbance. Globally after most intense disturbance (plots T2 and T3)
communities redundancy decreased at first place before increasing to
edge, recover or exceed the initial value.

@@

\caption{Trajectories of the difference from initial state of functional redundancy within the initial functional space over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}(\#fig:RedFun_rest)
\end{figure}

>>>>

From 1989 (2-3 years after disturbance) to 2015 (28-29
years after disturbance), 828388 individual trees and 591 botanical species
pertaining to 223 genus and 64 botanical families were recorded.
For undisturbed plots Richness and taxonomic evenness (Shannon and Simpson diversities) remained
stable over the 30 years monitored. In disturbed communities the
taxonomic richness increased after low disturbance intensity, reaching a
maximum gain of 14 botanical genera (plot 3 from treatment 2) while it
followed unimodal trajectories after intense disturbance, decreasing for
ten years before recovering to pre-disturbance values. In all disturbed
plots the taxonomic evenness (Shannon and Simpson diversities)
increased, following unimodal trajectories with a maximum, reached after
around 20 years, positively correlated to the disturbance treatment
(\(\rho_{spearman}^{Shannon}=0.86\), and
\(\rho_{spearman}^{Simpson}=0.89\)). Return towards initial evenness
values was beginning after 30 years except for two T3 plots (plots 8 and
12) which evenness still increased, suggesting similar but delayed
trajectories \ref{fig:IDHplot}.

@@

\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance. Colors are treaments: green (control), blue (T1), orange (T2), red (T3) with sahded areas the credibility intervals. Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities, 10, 20 and 30 years after disturbance}\label{fig:IDHplot}
\end{figure*}

@@

While both taxonomic and functional composition remained stable in
undisturbed communities (ref to fig), they followed consistent trajectories over time
after disturbance. According to the mapping of functional traits (see appendix I) these
compositional changes corresponded to shifts towards species with more
acquisitive functional strategies, from communities with high average WD
to high average SLA and chlorophyll content. For disturbed communities
the distance of successive inventories to the 1989 reference inventory
followed unimodal trajectories translating cyclic compositional changes
with a recovery of the initial composition (Figure \ref{fig:NMDSplans}).
The maximum dissimilarity with the initial state was positively
correlated to the disturbance treatment for both taxonomic and
functional composition (\(\rho_{spearman}^{taxonomic}=0.91\) and
\(\rho_{spearman}^{functional}=0.96\) respectively) and the time at
maximum was reached around 26 years after disturbance for taxonomic
composition and 22 years for functional composition.

@@

\caption{Plot trajectories in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) in a the two-dimensional NMDS space. Lower panels (\textbf{(c)} and \textbf{(d)}) represent the euclidean distance to initial condition along the 30 sampled years. Colors are disturbance treatments (green for control, blue for T1,orange for T2 and red for T3) with shaded areas, the credibility intervals}\label{fig:NMDSplans}
\end{figure*}

@@

Changes in functional composition trajectories went hand to hand
with consistent trajectories of the 8 functional traits (Figure
\ref{fig:CWM}). Except for leaf chlorophyll content, which continued to increase for
some T3 and T2 plots 30 years after disturbance, all traits and seed
mass proportions followed unimodal trajectories either stabilizing or
returning towards their initial values. Thirty years after disturbance
the weighted means of communities specific maximum height at adult stage
(\emph{Hmax}), leaf toughness (\emph{L\_toughness}) and wood specific
gravity (\emph{WD}) remained significantly lower than their initial
value (Figure \ref{fig:CWM}). The weighted means of bark thickness
(\emph{Bark\_thick}) similarly remained substantially higher than
initially for all disturbed plots while the specfic leaf area
(\emph{SLA}) had almost recovered its initial value. For all traits the
maximum difference to initial state was correlated to the disturbance
intensity (\(\rho_{spearman}^{L_{thickness}}=0.67\),
\(\rho_{spearman}^{L_{chloro}}=0.45\),
\(\rho_{spearman}^{L_{toughness}}=-0.43\),
\(\rho_{spearman}^{SLA}=0.93\), \(\rho_{spearman}^{WD}=-0.78\),
\(\rho_{spearman}^{Bark-thickness}=0.88\),
\(\rho_{spearman}^{Hmax}=-0.48\)).

@@

\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}). Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
\end{figure*}

@@

Communities functional redundancy remained stable in control
plots but after disturbance the redundancy trajectories were quite
variable (See appendix I) and apparently independently of the initial
disturbance. Globally after most intense disturbance (plots T2 and T3)
communities redundancy decreased at first place before increasing to
edge, recover or exceed the initial value. *I would suggest to consider only the Functional redundancy restricted to the functional space in the paper*

@@

\caption{Trajectories of functional redundancy within the initial functional space over 30 years after disturbance. Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}(\#fig:RedFun_rest)
\end{figure}
